### 2013-07-24 5/applications.tex
updated for Ian changes

\chapter{Using the Consistent Measures of Distance} %chapter 5
\label{chap:applications}

\section{Clustering SAR images using the consistent measures}
%\section{SAR Image Clustering using the consistent measures}

%\section{Proposed SAR speckle filter using Homoskedastic features}
\label{sec:k_mle}

Given an area of analysis, 
  the question of whether it is homogeneous or not can be investigated using
%its non-homogeneity can be detected by 
the statistical hypothesis test described above. 
If that is decided to be the case, clustering algorithm can be applied to partition the area into smaller regions. 
Thus an image can be segmented into homogeneous areas by applying the process repeatedly.

%Given an area of analysis, its non-homogeneity can be detected by the statistical hypothesis test described above. 
%If that is likely to be the case, clustering algorithm can be applied to partition the area into smaller regions. 
%The process can be applied iteratively until all partitions can be considered homogeneous (accepting the null hypothesis). 
%This forms the basis of our proposed speckle filtering technique.

\subsection{Developing kMLE clustering Algorithm}

@@

In a window of analysis, intensity dataset $X^I=\{x^I_1,...,x^I_n\}$ follows the exponential distribution given earlier. 
Assuming prior knowledge of it being homogeneous, thus having a single $\sigma$, the aim then is to estimate $\sigma$. 
This section will show how $\sigma$ can be estimated in log-transformed domain.

It has already been proven that the log-transformed dataset $X^l=\log_2 \sigma$ follows a Fisher-Tippet distribution $pdf(x)= \ln 2 \frac{ 2^d }{ e^{2^d} }$, with $d=x-\mu$, and $\mu=2 \log_2 \sigma$. 
Thus the problem of estimating $\sigma$, casting into the log-transformed domain, equals to the problem of estimating $\mu$. 
The PDF given in Eqn. \ref{eqn:conditional_prob} then can be interpreted as
  the conditional probability of observing a single value of $x^l_i$ when the underlying coefficient equals $\mu$: %is estimated, $\sigma$ can also be estimated $\sigma=2^{ \frac{\mu}{2} }$.

\begin{eqnarray}
\label{eqn:conditional_prob}
P(x^l_i | \mu ) 
	&=& \ln 2 \frac{ 2^{x^l_i - \mu} }{ e^{2^{ x^l_i -\mu }} }
%\displaystyle{ \frac{ \partial }{\partial{ \mu }} P(x^l_i | \mu )  }
%	&=& - \ln^2{2} \frac{ 2^{x^l_i - \mu} (2^{x^l_i - \mu}-1) }{ e^{2^{ x^l_i -\mu }} } 
\end{eqnarray}

Since log-transformed variables are biased, an MLE approach is used to estimate $\mu$. 
The MLE strategy aims to maximize the conditional (likelihood) probability of observing a set of value X found in the area, given the underlying coefficient equals $\mu$: $P(X^l | \mu )$. 
Assuming the values of $x^l_i$ are independent, the statistical product rule indicates that $P(X^l | \mu )=\displaystyle{\prod^n_{i=1}{P(x^l_i | \mu ) }}$, with  $P(x^l_i | \mu)$ as given in equation \ref{eqn:conditional_prob}.   
%$P(x^l_i | \mu ) = \ln 2 \frac{ 2^{x^l_i - \mu} }{ e^{2^{ x^l_i -\mu }} }$. 
Maximize $P(X^e | \mu )$ also maximizes $\ln[P(X^l | \mu )]$, which can be written as:

\begin{eqnarray}
\ln[P(X^l | \mu )] 
	= \displaystyle{\sum^n_{i=1}{ \ln(P(x^l_i | \mu )) }} 
	= \displaystyle{\sum^n_{i=1}{ \ln \left[ \ln 2 \frac{ 2^{x^l_i - \mu} }{ e^{2^{ x^l_i -\mu }} } \right] }} 
%	&=& n \ln \ln 2 +  \displaystyle{\sum^n_{i=1}{ \ln \left( 2^{x^l_i - \mu} \right) - \ln \left( e^{2^{ x^l_i -\mu }} \right) }} 
\end{eqnarray}

The parameter $\mu$ which maximizes the likelihood is found by setting its derivatives to zero:
\begin{eqnarray}
\displaystyle{ \frac{ \partial{ \ln[P( X^l | \mu )] }  }{\partial{ \mu }} }
	&=& 0 \\
\displaystyle{ \frac{ \partial }{\partial{ \mu }} \sum^n_{i=1}{ \ln( 2^{x^l_i - \mu}) - \ln( e^{2^{ x^l_i -\mu }} ) } }
	&=& 0 
%\displaystyle{ \frac{ \partial }{\partial{ \mu }} \sum^n_{i=1}{ \ln( 2^{x^l_i - \mu}) } }
%	&=& - n \cdot \ln2  \\
%\displaystyle{ \frac{ \partial }{\partial{ \mu }} \sum^n_{i=1}{ \ln( e^{2^{ x^l_i -\mu }} )i } }
%	&=& - \ln2 \cdot \sum^n_{i=1}{ 2^{ x^l_i -\mu } } \\
%	&=& - \ln2 \cdot \frac{ \sum^n_{i=1}{ 2^{ x^l_i } }}{2^\mu}  
\end{eqnarray}

Thus $\mu$ can be found by:
\begin{eqnarray}
%n \cdot \ln2	
%	&=& \ln2 \cdot \frac{ \sum^n_{i=1}{ 2^{ x^l_i } }}{2^\mu}  \\
2^\mu 
	&=& \frac{ \displaystyle{ \sum^n_{i=1}{ 2^{ x^l_i } }} }{n}  
\label{eqn:mle_log_transform}
\end{eqnarray}

@@

In order to apply MLE into a given analysis window ($L^w$), one would need to partition the area into groups of homogeneous pixels ($G$). 
Assuming for now that the number of homogeneous groups ($c=size(C)$) is known \textit{a priori}, this section will describe the proposed clustering algorithm. 
The basic template is similar to the k-means clustering algorithms (Listing \ref{lst:k_mle_algo_template})

@@

%The pseudo code for the algorithm is provided in Listing \ref{lst:k_mle_algo_cluster_centers}

@@

Due to the special characteristics of the log-transformed domain, we have adopted
an MLE estimator in lieu of a standard averaging estimation for cluster centers. 
Consequently the kMLE algorithm is presented in Listing \ref{lst:k_mle_algo_cluster_centers}.
%a standard k-means algorithm transformed into k-mle algorithm given in Listing \ref{lst:k_mle_algo_cluster_centers}.

\lstset{language=Matlab,caption=Iteratively find better cluster centers,label=lst:k_mle_algo_cluster_centers}
\definecolor{darkgray}{rgb}{0.95,0.95,0.95}
\lstset{backgroundcolor=\color{darkgray}}
%\begin{figure}
%\begin{verbatimtab}
\begin{lstlisting}[frame=tb]
function c_new = find_new_centers(l, c)
  for (i = 1 to size_of(l))
    for (j = 1 to size_of(c))
      m[i,j] = 0; %reset memberships
      p[i,j] = conditional_probability_of( l[i]|c[j] );
    end for

@@

  return c_new;
end function
\end{lstlisting}
%\end{verbatimtab}
%\caption{Iteratively find better cluster centers}
%\label{src:k_mle_algo}
%\end{figure}

When compared with standard k-means the following differences are evident:
\begin{enumerate}
  \item In k-means, data points are grouped into clusters with minimum distance. 
In our specific situation, as the dispersion PDF is non-symmetric around the cluster center, log-transformed data points are assigned to the cluster that gives them highest likelihood probability.
  \item The calculation of new cluster centers is different also. 
In k-means, new cluster centres are estimated as the average of all membered data points. 
Here, due to the bias of the PDF, maximum likelihood estimations are preferred.
  \item All the design decisions are made possible by the knowledge of the statistical PDF in the log-transformed domain!
\end{enumerate}

@@

In the previous section the algorithm was designed with the assumption that the number of clusters are known \textit{a priori}. 
In practice, this is not the case. This section will describe how the issue is to be tackled.

@@

\lstset{language=Matlab,caption=Homogenous partitions number growing,label=lst:k_mle_algo_partition_number}
\definecolor{darkgray}{rgb}{0.95,0.95,0.95}
\lstset{backgroundcolor=\color{darkgray}}
%\begin{figure}
%\begin{verbatimtab}
\begin{lstlisting}[frame=tb]
function centers = partition_into_homogenous_groups(l, threshold)
  c_init = maximum_likelyhood_of(l);
  [further_partition_needed, c_new] = 
          check_homogeneity_status(l, c_init, threshold);

@@

%\section{Experiment Results and Discussions}%describe analysis and compare
\subsection{Evaluating the kMLE algorithm}
\label{sec:k_mle_result}

\subsubsection{Using the kMLE clustering algorithm for speckle filtering}

It should be noted there is a difference between accepting the null hypothesis and simply failing to reject it. 
However, in the section that follows, the decision rule in use is that if the sample variance is lower than the cut-off value, the null hypothesis is accepted. 
The next analysis shows that for the purpose of speckle-filtering the difference is negligible.

The principle of speckle filtering is to group pixels into homogeneous areas so that the stochastics noise can be removed.
This allows the kMLE clustering algorithm to form the basis of a speckle filtering technique,
  where the results of the clustering algorithm can be transformed into answers for the problem of speckle filter.
Specifically the clustered areas directly serves as the grouping of pixels into homogeneous areas, and the cluster centers serves as the estimated backscattering coefficient for the analysis window.

Let us now consider some extreme threshold-values used as the homogeneous criteria for grouping pixels in the kMLE speckle filter.
If threshold-value is set to a very high value (close to 100\%), then a given partition would almost never be broken down.
The kMLE speckle filter effect would then equals to a box-car filter. 
Conversely, if the threshold-value is set to a very low value (close to 0\%), then a given partition would almost always be broken down to a single value.
The kMLE speckle filtering effect then would be virtually negligible.

Figure \ref{fig:k_mle_diff_threshold} shows various degrees of speckle filtering controllable by using different threshold-values.
Using the same moving window size, here $3\times3$, the effect of filtering is apparently proportional to the threshold values employed:
  with sub-figures \ref{fig:k_mle_threshold_original} to \ref{fig:boxcar_filter} visualise qualitatively the effects of the kMLE speckle filter when the threshold is set at 0\% (or left untouched), 25\%, 80\% and 100\% (or equals to the boxcar filter) respectively.

%	Fig. \ref{fig:k_mle_threshold_original} shows the original intensity image the speckle left untouched (0\%), 
%	Fig. \ref{fig:k_mle_threshold_25pc} and Fig. \ref{fig:k_mle_threshold_80pc} visualise qualitatively the effects of the k-MLE filter when the threshold is set at 25\% and 80\% respectively, 
%	while Fig. \ref{fig:boxcar_filter} shows that too much filtering (boxcar, i.e. 100\%) may lead to blurring effects.

@@

One way to improve the speckle suppression effect is by working on a larger window size. 
To reduce the associated blurring effect,
  a lower threshold-value than 100\% compared to the box-car filter would help
    by providing an adaptive partitioning of the analysis window.
Figure \ref{fig:k_mle_iterative_vs_boxcar} shows the visual performance of k-MLE filtering on $5 \times 5$ window.

@@

Another way to improve the raw kMLE speckle filter is to apply it iteratively.
As can be seen in figure \ref{fig:k_mle_diff_threshold}, a normal 3x3 k-mle filter would not filter as much noise as a similar boxcar filter would. 
At the same time, a 5x5 boxcar filter while having better speckle suppression power, would lose out by the blurring effects (see Fig. \ref{fig:boxcar_5x5}). 
To improve k-MLE filtering power while avoiding the blurring effect, the kMLE speckle filter can be used iteratively. 
Presented in the figure \ref{fig:k_mle_filtering_power} is the result of a k-MLE filter configured in a 3x3 window with 25\% threshold-value, being applied iteratively for four times. 
%The filtering power is illustrated and the plot shows similar performance with a 5x5 boxcar filter, while the filtered image is rendered with much less blurring effect.
From the plot, the $3 \times 3$ iterative k-MLE filter appears to have similar quantitative performance with the $5 \times 5$ boxcar filter,
  while its visual imagary output appears much sharper, qualititatively speaking.

@@

\subsubsection{Evaluation of the kMLE speckle filter on Simulated targets}

A number of point targets were simulated on a homogeneous background clutter. 
The ground truth SNR between the target and background clutter is fixed at 9dB. 
The original, speckled images are presented in Figure \ref{fig:sar_point_target_test_boxcar_k_mle}.
Subsequently, the kMLE speckle filter is to be benchmarked against the standard boxcar filters.
%Both qualitative and quantitative comparisons is done between boxcar filters and the proposed iterative k-MLE filter.
Qualitatively speaking, it is apparent that the kMLE speckle filter delivers
a resultant image with both sharper edges than the 5x5 boxcar filter and a stronger visual contrast  than the 3x3 boxcar filter.
%a sharper image than the 5x5 boxcar filter, at the same time its result has better contrast than the 3x3 boxcar filter.
Quantitatively, the k-MLE also appears to have the best histogram response, of all the 3 filters compared.

@@

\subsubsection{Evaluation of the kMLE speckle filter on Real Images}

The filters are applied on a SLC RadarSat 2 image covering the Muada Merbok area of Malaysia. 
Figure \ref{fig:k_mle_on_RADARSAT2_data} shows the patches of original and filtered images of both natural and urban landscape.
Apparently the kMLE speckle filter delivers a significantly ``better'' image than the original data.

@@

%SAR is stochastic and SAR speckle filtering is the removal of stochastic ``noise''.
The principle of many speckle filters in removing the stochastic noise is to group data into homogeneous areas.
In the boxcar filter, the filtering process can be described via Eqn. \ref{eqn:boxcar}, 
where 
	$I_i$ is the intensity value of pixel \textit{i} and
	$n$ is the total number of pixel in the surrounding areas.
The implicit assumption is that ALL pixels in the surrounding area belongs to the same homogeneous area with the estimating central point.
%Clearly, in real-life images, the assumption in boxcar filter is too easy to get violated.
%In a filter previously proposed by the authors
In the kMLE filter presented above and in \citep{Le_2010_ACRS}, the surrounding region is partitioned into homogeneous areas and the approximation is given in Eqn. \ref{eqn:kmle},
where 
	$I_i$ is the intensity value of pixel \textit{i} and
	$k$ is the total number of pixels in the partitioned and homogeneous area covering the centre point.

\begin{tabular}{c c}
\begin{minipage}[c]{0.45\textwidth}
\begin{equation}
\label{eqn:boxcar}
f_{boxcar} = \displaystyle{\sum_{i=1}^n{I_i}} / n 
\end{equation}
\end{minipage}
\begin{minipage}[c]{0.45\textwidth}
\begin{equation}
\label{eqn:kmle}
f_{kMLE }= \displaystyle{\sum_{i=1}^k{I_i} } / k 
\end{equation}
\end{minipage}
\end{tabular}

%By and large,
While it appears logical to classify surrounding pixels as being either in the same homogeneous area with the central point or not,
  the concept of ``homogeneous area'' lacks precise definition and boundaries.
%This is self-evident in the need for a threshold on the variance measured to assert homogeneity in an ensemble of stochastic samples.
In fact, the classification decisions between homogeneity and heterogeneity are probabilistic in nature.
%the concept of “homogeneous area” lacks precise definition and boundaries. 
This is evident in the statistical hypothesis testing used to support these decision makings.

The fMLE speckle filter presented in this section follow fuzzy logic approach,
  which, instead of fixing on a rigid two-member set of logical values (i.e. true and false),
  embraces truth values that range from 1 to 0.
Thus instead of making an YES/NO decision on whether or not a data point is in the same homogeneous region with the estimating central point, a fuzzy probabilistic scale is calculated.
Then the noise-removal impact of the surrounding data point will be scaled with this possibilistic probability,
  in comparison to the usual scheme that proceeds with ``full impact'' or ``no impact at all''.
The benefit of this fuzzy logic approach is that: we no longer need to worry about the homogeneous / heterogeneous problem.

%The idea presented in this paper is structured as follows:
%Section \ref{sec:fmle_estimation} will describe the FMLE estimation in details.
%The performance of the proposed filter will be evaluated in Section \ref{sec:results_eval}.
%Finally, Section \ref{sec:conclusion} concludes this paper.
%The last section will offers some conclusion and finalizes the paper.

\subsection{The FMLE SAR Speckle Filters}

The foundation of the fuzzy-estimation approach is the consistent sense of distance found in the log-transformed domain.
%It is shown in \cite{Le_2010_ACRS}
The consistent sense of distance states that given two SAR intensity samples coming from the same background, their distance in the log transformed domain follows a fixed distribution.

The base-2 log transformation is defined as:
${L_I}^i = \log_2{(I^i)}$ 
where 
	$I^i$ is the \textit{$i^{th}$} intensity sample in the originally measured domain,
	$L_I^I$ is the value of that sample in the log-transformed domain.
%Distance in log-transformed domain is defined as:
The log-contrast is then defined as:
$D= L_I^1 - L_I^2$. 
Assumming $I^i$ follows a negative exponential distribution \citep{Goodman_JOptSocAm_76}, 
  then this dis-similarity distance will follow the distribution 
$pdf(D)= \frac{2^D}{(1+2^D)^2} \ln2$,
  which is consistent.
This analysis is confirmed by real-life data validations (see \cite{Le_2010_ACRS}) and visually demonstrated in Fig. \ref{fig:consistent_contrast_log_domain}.

\subsubsection{FMLE Estimation}

In this fuzzy logic approach, a pixel with intensity value $I_i$ in the surrounding area of the center point $I_0$ has a likelihood of being in the same homogeneous area given as: 
$p_i = pdf(D_i)$, with $D_i=log_2(I_i) - log_2(I_0)$.
The fMLE approximation is then proposed as written in Eqn. \ref{eqn:fmle_incl}.

%We will see how this works in an homogeneous simulated area.
%The figures demonstrate speckle being filtering effects.
In implementing this statistical estimator, there are very noticable black dots in the filtered output image.
These dots can be explained as the long-tailed nature of the intensity PDF, 
  which manifests itself in the significant existence of very small values in the population.
In this all-inclusive scheme, there is 
	a single $D_0=0$ (at $I_0$) in the middle of other very large $D_{i, (i > 0)}$, which results in
	a single significant $p_0$ (at $I_0$), and other probably insignificant $p_{i, (i>0)}$.
This marginalises the filtering power of surrounding pixels.
One way to counter this effect is to exclude the centre point $I_0$ from its estimation,
  which results in the exclusive scheme written in Eqn. \ref{eqn:fmle_excl}.

>>>>

\chapter{Applications of The Proposed Models} %chapter 5
\label{chap:applications}

In this chapter, some beneficial applications of the proposed models are presented,
  which demonstrate how the models' consistent sense of variance capability can help to break the vicious circle encountered in SAR speckle
filtering .
This chapter also shows how the MSE can then be 
applicable in the additive and homoskedastic domain to evaluate SAR speckle filters, and further extends the concept to also evaluate
 POLSAR speckle filters.

\section{Clustering SAR images using the consistent measures}
\label{sec:k_mle}

Given an area to be analyzed, 
  the question of whether it is homogeneous or not can be investigated using
the statistical hypothesis test described in chapter \ref{chap:sar}. 
If that is found to be the case, clustering algorithm can then be applied to partition the area into smaller regions. 
Thus an image can be segmented into homogeneous areas by applying the process repeatedly.

\subsection{Developing k-MLE clustering Algorithm}

@@

In a window of analysis, intensity dataset $X^I=\{x^I_1,...,x^I_n\}$ follows the exponential distribution given in Eqn. \ref{eqn:SAR Intensity PDF}. 
Assuming prior knowledge of it being homogeneous, thus having a single $\sigma$, the objective then is to estimate $\sigma$. 
This section shows how $\sigma$ can be estimated in log-transformed domain.

It has already been proven that the log-transformed dataset $X^l=\log_2 \sigma$ follows a Fisher-Tippet distribution $pdf(x)= \ln 2 \frac{ 2^d }{ e^{2^d} }$, with $d=x-\mu$, and $\mu=2 \log_2 \sigma$. 
Thus the problem of estimating $\sigma$, casting into the log-transformed domain, equals to the problem of estimating $\mu$. 
The PDF given in Eqn. \ref{eqn:conditional_prob} can then be interpreted as
  the conditional probability of observing a single value of $x^l_i$ when the underlying coefficient equals $\mu$: %is estimated, $\sigma$ can also be estimated $\sigma=2^{ \frac{\mu}{2} }$.

%\begin{eqnarray}
%\displaystyle{ \frac{ \partial }{\partial{ \mu }} P(x^l_i | \mu )  }
%	&=& - \ln^2{2} \frac{ 2^{x^l_i - \mu} (2^{x^l_i - \mu}-1) }{ e^{2^{ x^l_i -\mu }} } 
%\end{eqnarray}
\begin{equation}
\label{eqn:conditional_prob}
\caption{eqn:Conditional Probability of SAR Pixel Intensity}
P(x^l_i | \mu ) 
	= \ln 2 \frac{ 2^{x^l_i - \mu} }{ e^{2^{ x^l_i -\mu }} }
\end{equation}

Since log-transformed variables are biased, an MLE approach is used to estimate $\mu$. 
The MLE strategy aims to maximize the conditional (likelihood) probability of observing a set of value X found in the area, given the underlying coefficient equals $\mu$: $P(X^l | \mu )$. 
Assuming the values of $x^l_i$ are independent, the statistical product rule indicates that $P(X^l | \mu )=\displaystyle{\prod^n_{i=1}{P(x^l_i | \mu ) }}$, with  $P(x^l_i | \mu)$ as given in Eqn. \ref{eqn:conditional_prob}.   
%$P(x^l_i | \mu ) = \ln 2 \frac{ 2^{x^l_i - \mu} }{ e^{2^{ x^l_i -\mu }} }$. 
Maximize $P(X^e | \mu )$ also maximizes $\ln[P(X^l | \mu )]$, which can be written as:

\begin{eqnarray*}
\ln[P(X^l | \mu )] 
	= \displaystyle{\sum^n_{i=1}{ \ln(P(x^l_i | \mu )) }} 
	= \displaystyle{\sum^n_{i=1}{ \ln \left[ \ln 2 \frac{ 2^{x^l_i - \mu} }{ e^{2^{ x^l_i -\mu }} } \right] }} 
%	&=& n \ln \ln 2 +  \displaystyle{\sum^n_{i=1}{ \ln \left( 2^{x^l_i - \mu} \right) - \ln \left( e^{2^{ x^l_i -\mu }} \right) }} 
\end{eqnarray*}

The parameter $\mu$ which maximizes the likelihood is found by setting its derivatives to zero:
\begin{eqnarray*}
\displaystyle{ \frac{ \partial{ \ln[P( X^l | \mu )] }  }{\partial{ \mu }} }
	&=& 0 \\
\displaystyle{ \frac{ \partial }{\partial{ \mu }} \sum^n_{i=1}{ \ln( 2^{x^l_i - \mu}) - \ln( e^{2^{ x^l_i -\mu }} ) } }
	&=& 0 
%\displaystyle{ \frac{ \partial }{\partial{ \mu }} \sum^n_{i=1}{ \ln( 2^{x^l_i - \mu}) } }
%	&=& - n \cdot \ln2  \\
%\displaystyle{ \frac{ \partial }{\partial{ \mu }} \sum^n_{i=1}{ \ln( e^{2^{ x^l_i -\mu }} )i } }
%	&=& - \ln2 \cdot \sum^n_{i=1}{ 2^{ x^l_i -\mu } } \\
%	&=& - \ln2 \cdot \frac{ \sum^n_{i=1}{ 2^{ x^l_i } }}{2^\mu}  
\end{eqnarray*}

Thus $\mu$ can be found by:
\begin{equation}
2^\mu 
	= \frac{ \displaystyle{ \sum^n_{i=1}{ 2^{ x^l_i } }} }{n}  
\label{eqn:mle_log_transform}
\caption{eqn:MLE Estimation in the Log-Transformed Domain}  
\end{equation}

@@

In order to apply MLE into a given heterogeneous analysis window ($L^w$), one would need to partition the area into groups of homogeneous pixels ($G$). 
Assuming for now that the number of homogeneous groups ($c=size(C)$) is known \textit{a priori}, this section will describe the proposed clustering algorithm. 
The basic template is similar to the k-means clustering algorithms (Listing \ref{lst:k_mle_algo_template})

@@

@@

Due to the special characteristics of the log-transformed domain, 
an MLE estimator is used in lieu of the standard averaging estimation for cluster centers. 
Consequently the k-MLE algorithm is presented in Listing \ref{lst:k_mle_algo_cluster_centers}.

\lstset{language=Matlab,caption=Iteratively find better cluster centers,label=lst:k_mle_algo_cluster_centers}
\definecolor{darkgray}{rgb}{0.95,0.95,0.95}
\lstset{backgroundcolor=\color{darkgray}}
\begin{lstlisting}[frame=tb]
function c_new = find_new_centers(l, c)
  for (i = 1 to size_of(l))
    for (j = 1 to size_of(c))
      m[i,j] = 0; %reset memberships
      p[i,j] = conditional_probability_of( l[i]|c[j] );
    end for

@@

  return c_new;
end function
\end{lstlisting}

Compared with using the standard k-means algorithm, the following differences are noted:
%When compared with standard k-means the following differences are evident:
\begin{enumerate}
  \item In k-means, data points are grouped into clusters with minimum distance. 
In our specific situation, as the dispersion PDF is non-symmetric around the cluster center, log-transformed data points are assigned to the cluster that gives them highest likelihood probability.
  \item The calculation of new cluster centers is also different. 
In k-means, new cluster centres are estimated as the average of all membered data points. 
Here, due to the bias of the PDF, maximum likelihood estimations are preferred.
  \item All the design decisions are made possible by the knowledge of the statistical PDF in the log-transformed domain.
\end{enumerate}

@@

In the last section the algorithm was designed with the assumption that the number of clusters are known \textit{a priori}. 
In practice, this is not the case. This section will describe how the issue is to be tackled.

@@

\lstset{language=Matlab,caption=Homogenous partitions number growing,label=lst:k_mle_algo_partition_number}
\definecolor{darkgray}{rgb}{0.95,0.95,0.95}
\lstset{backgroundcolor=\color{darkgray}}
\begin{lstlisting}[frame=tb]
function centers = partition_into_homogenous_groups(l, threshold)
  c_init = maximum_likelyhood_of(l);
  [further_partition_needed, c_new] = 
          check_homogeneity_status(l, c_init, threshold);

@@

%\section{Experiment Results and Discussions}%describe analysis and compare
\subsection{Evaluating the k-MLE algorithm}
\label{sec:k_mle_result}

\subsubsection{Using the k-MLE clustering algorithm for speckle filtering}

In the context of statistical hypothesis testing, there is a difference between accepting the null hypothesis and simply failing to reject it. 
However, in the section that follows, the decision rule in use is that if the sample variance is lower than the cut-off value, the null hypothesis is accepted. 
This section illustrates that for the purpose of speckle-filtering the difference is negligible.

The principle of speckle filtering is to group pixels into homogeneous areas so that the stochastics noise can be removed.
Hence the k-MLE clustering algorithm can be used to form the basis of a speckle filtering technique.
Specifically the clustered areas directly serves as the grouping of pixels into homogeneous areas, and
the cluster centers serves as the estimated backscattering coefficient for the analysis window.

Considering the case where some extreme threshold-values are used as the homogeneous criteria
If threshold-value is set to a very high value (close to 100\%), then a given partition would almost never be broken down.
The kMLE speckle filter effect would then equals to a box-car filter. 
Conversely, if the threshold-value is set to a very low value (close to 0\%), then a given partition would almost always be broken down to a single value.
The kMLE speckle filtering effect then would be virtually negligible.

Figure \ref{fig:k_mle_diff_threshold} shows various degrees of speckle filtering controllable by using different threshold-values.
Using the same moving window size of $3\times3$, the effect of filtering is clearly proportional to the threshold values employed.
This can be observed from Figs. \ref{fig:k_mle_threshold_original} to \ref{fig:boxcar_filter}
where the  visual quality  reflects the effects of the kMLE speckle filter when its  
threshold is set at 0\% (i.e. left untouched), 25\%, 80\% and 100\% (i.e. equals to the boxcar filter) respectively.

@@

One way to improve the speckle filtering  performance is to use a larger window size.
Fig. \ref{fig:boxcar_5x5} shows the visual quality performance of k-MLE filtering on $5 \times 5$ window.
As can be seen, the blurring effect in comparison to $3 \times 3$ filters is clearly visible.
To reduce the associated blurring effect, a threshold-value lower than 100\%  would help by providing an adaptive partitioning of the analysis
window (via the k-MLE clustering algorithm).

@@

Another way to improve the raw k-MLE speckle filter is to apply it iteratively.
As can be seen in Fig. \ref{fig:k_mle_diff_threshold}, a normal 3x3 k-MLE filter would not filter as much noise as a 3x3 boxcar filter. 
At the same time, a 5x5 boxcar filter while having better speckle suppression power, would lose out by its blurring effects as can be seen in Fig. \ref{fig:boxcar_5x5}. 
To improve k-MLE filtering power while avoiding the blurring effect, the k-MLE speckle filter can be used iteratively. 
Presented in the figure \ref{fig:k_mle_filtering_power} is the result of a k-MLE filter configured in a 3x3 window with 25\% threshold-value, being applied iteratively for four times. 
From the plot, the $3 \times 3$ iterative k-MLE filter appears to have similar quantitative performance with the $5 \times 5$ boxcar filter,
  while its visual imagery output appears much sharper, qualitatively speaking.

@@

\subsubsection{Evaluation of the kMLE speckle filter on simulated targets}

A number of point targets were simulated on a homogeneous background clutter. 
In this test, the SNR between the target and background clutter is fixed at 9dB. 
%The original, speckled images are presented in Figure \ref{fig:sar_point_target_test_boxcar_k_mle}.
Subsequently, the kMLE speckle filter is to be benchmarked against the standard boxcar filters as shown in Fig. \ref{fig:sar_point_target_test_boxcar_k_mle}.
%Both qualitative and quantitative comparisons is done between boxcar filters and the proposed iterative k-MLE filter.
Qualitatively speaking, it is clear that the kMLE speckle filter delivers
a resultant image with both sharper edges than the 5x5 boxcar filter and a stronger visual contrast  than the 3x3 boxcar filter.
%a sharper image than the 5x5 boxcar filter, at the same time its result has better contrast than the 3x3 boxcar filter.
Quantitatively, the k-MLE also appears to have the best histogram response among the 3 filters compared.

@@

\subsubsection{Evaluation of the k-MLE speckle filter on Real Images}

The k-MLE filters are also applied on a SLC RadarSat 2 image covering the Muada Merbok area of Malaysia. 
Figure \ref{fig:k_mle_on_RADARSAT2_data} shows the patches of original and filtered images of both natural and urban landscape.
It is obvious that the kMLE speckle filter delivers a significantly ``better'' image than the original data.

@@

As mentioned earlier, speckle filters removes the stochastic noise by grouping data into homogeneous areas. 
In the boxcar filter, the filtering process can be described via Eqn. \ref{eqn:boxcar}, 
where 
	$I_i$ is the intensity value of pixel \textit{i} and
	$n$ is the total number of pixel in the surrounding areas.
The implicit assumption for this filter is that ALL pixels in the surrounding area belongs to the same homogeneous area with the estimating central point.
In the k-MLE filter presented in the previous section and in \citep{Le_2010_ACRS}, the surrounding region is partitioned into homogeneous areas and the approximation is based on Eqn. \ref{eqn:kmle},
where 
	$I_i$ is the intensity value of pixel \textit{i} and
	$k$ is the total number of pixels in the partitioned and homogeneous area covering the centre point.

%\begin{tabular}{c c}
%\begin{minipage}[c]{0.45\textwidth}
\begin{equation}
\label{eqn:boxcar}
\caption{eqn:Boxcar Speckle Filter Estimation}
f_{boxcar} = \displaystyle{\sum_{i=1}^n{I_i}} / n 
\end{equation}
%\end{minipage}
%\begin{minipage}[c]{0.45\textwidth}
\begin{equation}
\label{eqn:kmle}
\caption{eqn:k-MLE Speckle Filter Estimation}
f_{kMLE }= \displaystyle{\sum_{i=1}^k{I_i} } / k 
\end{equation}
%\end{minipage}
%\end{tabular}

While it appears logical to classify surrounding pixels as falling either inside or outside the same homogeneous area with the central point,
  the concept of ``homogeneous area'' lacks precise definition and boundaries.
In fact, the classification decisions between homogeneity and heterogeneity are probabilistic in nature.
This is evident in the statistical hypothesis testing used to support these decision makings \cite{Le_2010_ACRS}.

%The f-MLE speckle filter presented
In this section fuzzy logic approach is followed
  where, instead of fixing on a rigid two-member set of logical values (i.e. true and false),
  embraces truth values that range from 1 to 0.
Thus instead of making a YES/NO decision on whether or not a data point is in the same homogeneous region with the estimated central point, a fuzzy probabilistic scale is calculated.
Then the noise-removal impact of the surrounding data point will be scaled with this probability value.
This fuzzy logic approach give rise to the f-MLE speckle filter presented next,
  and its main benefit is that:  we no longer need to worry about the homogeneous / heterogeneous problem.

\subsection{The F-MLE SAR Speckle Filters}

The foundation of the fuzzy-estimation approach is the consistent sense of distance found in the log-transformed domain.
%It is shown in \cite{Le_2010_ACRS}
The consistent measure of distance states that given two SAR intensity samples coming from the same background, their distance in the log transformed domain follows a fixed distribution according to Eqn. \ref{eqn:SAR Log Intensity Contrast PDF}.

The base-2 log transformation is defined as:
${L_I}^i = \log_2{(I^i)}$ 
where 
	$I^i$ is the \textit{$i^{th}$} intensity sample in the originally measured domain,
	$L_I^i$ is the value of that sample in the log-transformed domain.
%Distance in log-transformed domain is defined as:
The contrast is then defined as:
$C= L_I^1 - L_I^2$. 
Assuming $I^i$ follows a negative exponential distribution \citep{Goodman_JOptSocAm_76}, 
  then this dis-similarity distance will follow the distribution 
$pdf(C)= \frac{2^C}{(1+2^C)^2} \ln2$,
  which is consistent.
This analysis is confirmed by real-life data validations (see \cite{Le_2010_ACRS}) and visually demonstrated in Fig. \ref{fig:consistent_contrast_log_domain}.

\subsubsection{F-MLE Estimation}

In this fuzzy logic approach, a pixel with intensity value $I_i$ in the surrounding area of the center point $I_0$ has a likelihood of being in the same homogeneous area given as: 
$p_i = pdf(C_i)$, with $C_i=log_2(I_i) - log_2(I_0)$.
The fMLE approximation is then proposed as written in Eqn. \ref{eqn:fmle_incl}.

%We will see how this works in an homogeneous simulated area.
%The figures demonstrate speckle being filtering effects.
In implementing this statistical estimator, there are very noticeable black dots in the filtered output image.
These dots can be explained as the long-tailed nature of the intensity PDF, 
  which manifests itself in the significant existence of very small values in the population.
In this all-inclusive scheme, there is 
	a single $C_0=0$ (at $I_0$) in the middle of other very large $C_{i, (i > 0)}$, which results in
	a single significant $p_0$ (at $I_0$), and other probably insignificant $p_{i, (i>0)}$.
This marginalizes the filtering power of surrounding pixels.
One way to counter this effect is to exclude the centre point $I_0$ from its estimation,
  which results in the exclusive scheme written in Eqn. \ref{eqn:fmle_excl}.

%\begin{tabular}{c c}
%\begin{minipage}[c]{0.45\textwidth}
\begin{equation}
\label{eqn:fmle_incl}
\caption{eqn:F-MLE Estimation (Inclusive Version)}
FMLE_{incl}= \frac{\displaystyle{\sum_{i=0}^n{p_i * I_i} } }{\displaystyle{\sum_{i=0}^n{p_i } } }
\end{equation}
%\end{minipage}
%\begin{minipage}[c]{0.45\textwidth}
\begin{equation}
\label{eqn:fmle_excl}
\caption{eqn:F-MLE Estimation (Exclusive Version)}
FMLE_{excl}= \frac{\displaystyle{\sum_{i=1}^n{p_i * I_i} } }{\displaystyle{\sum_{i=1}^n{p_i } } } 
\end{equation}
%\end{minipage}
%\end{tabular}
